# pre-class


Make sure you commit this often with meaningful messages. 




### Question 1:

Using a loop, print the integers from 1 to 50. 
```{r}

```
### Question 2:

A.  Using a loop, add all the integers between 0 and 1000.
```{r}
sum <- 0
for (i in 0:1000) {
  sum <- sum + i
}
# The sum of all the integers between 0 and 1000 is 500500.
```
B. Now, add all the EVEN integers between 0 and 1000 (hint: use seq())
```{r}
sum(seq(from = 0, to = 1000, by =2))

# The sum of all the even integers between 0 and 1000 is 250500.
```
C. Now, repeat A and B WITHOUT using a loop.
```{r}
All_Integers <- c(0:1000)
sum1 <- sum(All_Integers)

Even_Integers <- seq(from = 0, to = 1000, by =2)
sum2 <- sum(Even_Integers)

# These produce the same results. Part B is the same because I didn't use a loop the first time.
```
### Question 3:

Here is a dataframe of survey data containing 5 questions :

```{r, eval=FALSE}
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )
```
The response to each question should be an integer between 1 and 5. Obviously, we have some bad values in the dataframe. The goal of this problem is to fix them.

A. Using a loop, create a new dataframe called survey.clean where all the invalid values (those that are not integers between 1 and 5) are set to NA.
```{r}
survey.clean <- data.frame()
for (i in 1:nrow(survey)) {
  for(j in 1:ncol(survey)) {
    if(survey[i,j]>=1 & survey[i,j]<=5) {
      survey.clean[i,j] = survey[i,j]
    } else {
      survey.clean[i,j] = NA
    }
  }
}
```

B. Now, again using a loop, add a new column to the dataframe called “invalid.answers” that indicates, for each participant, how many bad answers they gave.
```{r}


```

