# pre-class


Make sure you commit this often with meaningfull messages. 


### Getting Started


We will work with the dataset called [gapminder](https://github.com/jennybc/gapminder), this is a cleaned up version from [Gapminder Data](http://www.gapminder.org/data/). Gapminder contains a lot of great data on all of the nations of the world. We first need to install the gapminder package in R. 

```{r}
#install.packages("gapminder")
```



```{r}
library(dplyr)
library(gapminder)
gapminder
```



### Pre-Class Problems

Use **dplyr** functions to address the following questions:

1. How many unique countries are represented per continent?

```{r}
CountryDataTable <- as.tbl(gapminder) 

Unique_Countries <- CountryDataTable%>%
  group_by(continent) %>%
    distinct(country) %>%
      tally() %>%
        arrange(desc(by = n))

Unique_Countries

# Africa has 52 countries, Asia has 33 countries, Europe has 30 countries, Americas has 25 countries, and Oceania has 2 countries.
```

2. Which European nation had the lowest GDP per capita in 1997? 

```{r}
European_LowGDP <- CountryDataTable %>%
  filter(continent == "Europe", year == 1997) %>% # Filtered by country and year (Europe, 1997)
    arrange(gdpPercap) %>% # Arranged from least to greatest by GDP per Capita
      select(country, gdpPercap) %>%
        slice(1)

European_LowGDP
# The European nation with the lowest GDP per capita in 1997 was Albania, with a value of 3193.       
```



3. According to the data available, what was the average life expectancy across each continent in the 1980s?

```{r}
Avg_LE <- CountryDataTable %>%
  group_by(continent) %>%
    filter(year>=1980 & year<=1989) %>%
      summarize(Avg_lifeExp = mean(lifeExp)) %>%
       arrange(Avg_lifeExp)

Avg_LE 

# I first grouped the data by continent, then calculated the mean life expectancy for each continent. After, I sorted from least to greatest.

# The average life expectancy was 48.9 for Africa, 60.1 for Asia, 64.7 for Americas, 71.9 for Europe, and 74.3 for Oceania. 
```

4. What 5 countries have the highest total GDP over all years combined?

```{r}
GDP_Total <- CountryDataTable %>%
  mutate(GDPTotal = gdpPercap*pop) %>%
    group_by(country) %>%
      summarize(GDP_Sum = sum(GDPTotal)) %>%
        arrange(desc(by = GDP_Sum)) %>%
          slice(1:5)

GDP_Total

# The United States, Japan, China, Germany, and the United Kingdom have the highest total GDP over all years combined.
```

5. What countries and years had life expectancies of _at least_ 80 years? _N.b. only output the columns of interest: country, life expectancy and year (in that order)._

```{r}
Expectancy <- CountryDataTable %>%
  filter(lifeExp >= 80) %>%
    select(country, lifeExp, year) %>%
      group_by(country)

Expectancy

# Most countries that had life expectancies of at leat 80 years were after 2000.
```

6. What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?

```{r}
Country_Correlation <- CountryDataTable %>%
  group_by(country) %>% # Grouped by country
   summarise(corr = abs(cor(lifeExp, gdpPercap))) %>% # calculated the absolute value of the correlation between life exp and GDP Per Cap
    arrange(desc(corr)) %>% #sorted from greatest to least
      slice(1:10)

Country_Correlation

# France has the strongest correlation between life expectancy and per capita GDP followed by Austria, Belgium, Norway, Oman, United Kingdom, Italy, Israel, Denmark, and Australia.
```

7. Which combinations of continent (besides Asia) and year have the highest average population across all countries? _N.b. your output should include all results sorted by highest average population_. With what you already know, this one may stump you. See [this Q&A](http://stackoverflow.com/q/27207963/654296) for how to `ungroup` before `arrange`ing. This also [behaves differently in more recent versions of dplyr](https://github.com/hadley/dplyr/releases/tag/v0.5.0).

```{r}
Highest_Avg_Pop <- CountryDataTable %>%
  filter(continent != "Asia") %>%
    group_by(continent, year) %>%
      summarise(avg_pop = mean(pop)) %>%
        arrange(desc(avg_pop))

Highest_Avg_Pop

# The Americas claim the top 8 spots, having the highest average population across all countries than the other continents.
```

8. Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data?

```{r}
Country_SD <- CountryDataTable %>%
  group_by(country) %>%
    summarise(SD = sd(pop)) %>%
      arrange(SD) %>%
        slice(1:3)

Country_SD

# Sao Tome and Principe, Iceland, and Montenegro had the most consistent population estimates across the years of available data.
```

9. Subset **gm** to only include observations from 1992 and store the results as **gm1992**. What kind of object is this?
```{r}
gm1992 <- subset(CountryDataTable, year == 1992)
typeof(gm1992)

# This object is a list.
```

10. Which observations indicate that the population of a country has *decreased* from the previous year **and** the life expectancy has *increased* from the previous year? See [the vignette on window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html).
```{r}
PopCountry <- CountryDataTable %>%
  group_by(country) %>%
   mutate(Population_Difference = pop - lag(pop), LifeExpectancy_Difference = lifeExp - lag(lifeExp)) %>%
    filter(Population_Difference < 0 & LifeExpectancy_Difference > 0) %>%
      select(country, year, lifeExp, pop)

PopCountry

# The observations in PopCountry show the population of a contry that decreased from the previous year but also had their life expectancy increase from the previous year.
```
