# pre-class


Make sure you commit this often with meaningfull messages. 


### Getting Started


We will work with the dataset called [gapminder](https://github.com/jennybc/gapminder), this is a cleaned up version from [Gapminder Data](http://www.gapminder.org/data/). Gapminder contains a lot of great data on all of the nations of the world. We first need to install the gapminder package in R. 

```{r}
install.packages("gapminder")
```



```{r}
library(dplyr)
library(gapminder)
gapminder
```



### Pre-Class Problems

Use **dplyr** functions to address the following questions:

1. How many unique countries are represented per continent?

```{r}
CountryDataTable <- as.tbl(gapminder) 

Unique_Countries <- CountryDataTable%>%
  group_by(continent) %>%
    distinct(country) %>%
      tally() %>%
        arrange(desc(by = n))

# Africa has 52 countries, Asia has 33 countries, Europe has 30 countries, Americas has 25 countries, and Oceania has 2 countries.
```

2. Which European nation had the lowest GDP per capita in 1997? 

```{r}
European_LowGDP <- CountryDataTable %>%
  filter(continent == "Europe", year == 1997) %>% # Filtered by country and year (Europe, 1997)
    arrange(gdpPercap) # Arranged from least to greatest by GDP per Capita
      
European_LowGDP[1,]

# The European nation with the lowest GDP per capita in 1997 was Albania, with a value of 3193.       
```



3. According to the data available, what was the average life expectancy across each continent in the 1980s?
```{r}
Avg_LE <- CountryDataTable %>%
  group_by(continent) %>%
   summarize(Avg_lifeExp = mean(lifeExp)) %>%
    arrange(Avg_lifeExp)

Avg_LE 

# I first grouped the data by continent, then calculated the mean life expectancy for each continent. After, I sorted from least to greatest.

# The average life expectancy was 48.9 for Africa, 60.1 for Asia, 64.7 for Americas, 71.9 for Europe, and 74.3 for Oceania. 
```



4. What 5 countries have the highest total GDP over all years combined?




5. What countries and years had life expectancies of _at least_ 80 years? _N.b. only output the columns of interest: country, life expectancy and year (in that order)._




6. What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?




7. Which combinations of continent (besides Asia) and year have the highest average population across all countries? _N.b. your output should include all results sorted by highest average population_. With what you already know, this one may stump you. See [this Q&A](http://stackoverflow.com/q/27207963/654296) for how to `ungroup` before `arrange`ing. This also [behaves differently in more recent versions of dplyr](https://github.com/hadley/dplyr/releases/tag/v0.5.0).




8. Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data? 



9. Subset **gm** to only include observations from 1992 and store the results as **gm1992**. What kind of object is this?



10. Which observations indicate that the population of a country has *decreased* from the previous year **and** the life expectancy has *increased* from the previous year? See [the vignette on window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html).
