# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}
f1 <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
# This function checks to see if the string given contains the prefix. If the string has the specified prefix, TRUE is returned. If not, FALSE is returned.
# Better Names: Prefix_Checker, Has_Prefix, Prefix


f2 <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
# This function takes a vector and removes the last element. The function returns the vector with the last element removed. If the vector has only one value, NULL is returned.
# Better Names: Remove_Last_Element, Remove_Last

f3 <- function(x, y) {
  rep(y, length.out = length(x))
}
# This function takes the length of the first argument, regardless of the value, and replicates the second value that many times.
# Better Names: Replicate, Replicate_Nums

```

2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?
```{r}
rnorm(10, mean = 0, sd = 1)
MASS::mvrnorm(n=10, mu = 0, Sigma = 1)

# The output for rnorm is a vector while the output of MASS::mvrnorm is a matrix. MASS::mvrnorm is more powerful than rnorm because it can produce more than one sample from a specified multivariate normal distribution. MASS::mvrnorm can do everything rnorm can do but it opens up more possibilities if you need more variables and tests. One way to make them more consistent is to have either MASS::mvrnorm return a vector or have rnorm return a matrix with 1 column.
```
3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 
```{r}
lapply(mtcars, function(x) {sd(x)/mean(x)})

```
4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.
```{r}
vapply(mtcars, sd, numeric(1))
```    
    b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)
```{r}
vapply(iris[vapply(iris, is.numeric, logical(1))], sd, numeric(1))
# This uses vapply twice since it checks to see if the column is numeric first, returning a vector of T and F. Then the columns that are T are passed into the outer vapply.
```